Multidimensional signals and analytic flexibility: Estimating degrees of freedom in human speech analyses

### 2022-09-19 RR_manuscript/Draft_RR.Rmd
update table, added section at the end of statistical analyses re. number of teams claiming an effect

# review ratings
review_mean <- msa_models %>% 
  summarise(mean_prating = mean(phon_rating, na.rm = TRUE),
            sd_prating = sd(phon_rating, na.rm = TRUE),
            mean_srating = mean(stat_rating, na.rm = TRUE),
            sd_srating = sd(stat_rating, na.rm = TRUE))
```

@@

\phantom{.}

```{r}
#| label: team-descriptives-table
#| echo: false
#| results: 'asis'
#| eval: true

# Team descriptives
tbl_team_desc <- msa_models %>% 
  select(animal, years_from_phd, prior_belief, phon_rating, stat_rating, 
         all_rating) %>% 
  group_by(animal) %>% 
  summarize(across(years_from_phd:all_rating, mean)) %>% 
  pivot_longer(cols = -animal, names_to = "metric", values_to = "vals") %>% 
  na.omit() %>% 
  group_by(metric) %>% 
  summarize(min = min(vals), max = max(vals), avg = mean(vals), sd = sd(vals)) %>% 
  mutate_if(is.numeric, round, digits = 2) %>% 
  mutate(desc = "team_desc", 
    metric = fct_relevel(metric, "years_from_phd"), 
   `Mean (SD)` = glue::glue("{avg} ({sd})"), 
    Metric = case_when(
      metric == "years_from_phd" ~ "Years since PhD", 
      metric == "stat_rating" ~ "Statistical ability", 
      metric == "prior_belief" ~ "Prior belief of $H_{1}$", 
      metric == "phon_rating" ~ "Phon rating", 
      metric == "all_rating" ~ "all rating"), 
    label = glue::glue("{desc}_{metric}")) %>% 
  select(label, desc, Metric, Min. = min, Max = max, `Mean (SD)`, avg, sd) %>% 
  add_row(label = " ", desc = "team_desc", Metric = " ", 
          Min. = NA, Max = NA, `Mean (SD)` = NA, avg = NA, sd = NA) %>% 
  mutate(across(-c("label", "desc", "Metric"), ~replace(., is.na(.), " ")))

# Model descriptives 
tbl_model_desc <- msa_models %>% 
  select(animal, n_predictors:n_random, n_models) %>% 
  group_by(animal) %>% 
  summarize(across(everything(), mean)) %>% 
  pivot_longer(cols = -animal, names_to = "metric", values_to = "vals") %>% 
  na.omit() %>% 
  group_by(metric) %>% 
  summarize(min = min(vals), max = max(vals), avg = mean(vals), sd = sd(vals)) %>% 
  mutate_if(is.numeric, round, digits = 2) %>% 
  mutate(desc = "model_desc", 
    metric = fct_relevel(metric, "n_models"), 
    `Mean (SD)` = glue::glue("{avg} ({sd})"), 
    Metric = case_when(
      metric == "n_coefficients" ~ "n. coefficients", 
      metric == "n_models" ~ "n. models", 
      metric == "n_predictors" ~ "n. predictors", 
      metric == "n_random" ~ "n. random effects", 
      metric == "n_random_int" ~ "n. random intercepts", 
      metric == "n_random_slo" ~ "n. random slopes", 
      metric == "n_random_smooth" ~ "n. random smooths",
      metric == "n_terms" ~ "n. terms"), 
    label = glue::glue("{desc}_{metric}")) %>% 
  select(label, desc, Metric, Min. = min, Max = max, `Mean (SD)`, avg, sd) %>% 
  mutate(across(-c("desc", "Metric"), as.character))

# Combine in single tibble (for printing in prose and table)
tbl_team_model_desc <- bind_rows(tbl_team_desc, tbl_model_desc) %>% 
  mutate(desc = case_when(
    Metric != "all rating" & desc == "team_desc" ~ " ", 
    Metric != "n. coefficients" & desc == "model_desc" ~ " ", 
    TRUE ~ desc), 
    desc = case_when(
      desc == "team_desc" ~ "Teams", 
      desc == "model_desc" ~ "Models", 
      TRUE ~ " "))

tbl_team_model_desc %>% 
  select(Descriptives = desc, Metric, Min., Max, `Mean (SD)`) %>% 
  knitr::kable(format = "latex", booktabs = T, escape = F, 
    linesep = "", align = c("l", "l", "r", "r", "r"))

```

`r team_desc_list$Initial$n_teams$val %>% xfun::numbers_to_words() %>% stringr::str_to_sentence()` teams initially signed up to participate in the study.
`r (team_desc_list$Initial$n_teams$val - team_desc_list$Final$n_teams$val) %>% xfun::numbers_to_words() %>% stringr::str_to_sentence()` of the signed-up teams dropped out during the analysis phase.
One Team Coordinator, Wim van Dommelen (who agreed to their feedback being shared), gave the following reason: "The reason is that [...] the project's approach is too much top-down instead of bottom-up. [...] I would rather prefer to explore the collected utterances and run different types of acoustic analyses, which would then be followed by statistical evaluation of the data."
We find this contribution to clearly speak to the main objective of the current study: investigate researcher's degrees of freedom in the speech sciences.
Based on our personal experience with research in the field, it is common practice to test as many different types of models, using as many different types of measurements, to answer one research question.
The remaining `r team_desc_list$Final$n_teams$val` teams submitted their analysis by the established deadline.
These are the teams that are included in this study.

@@

@@

Based on statistical framework, type of model, distribution family, fixed terms, and not including random effects, there were a total of `r nrow(unique_model_specs)` different model specifications.
When considering both acoustic and statistical analysis, we have found a grand total of `r nrow(unique_acou_stat)`, i.e. each submitted analysis was unique.

>>>>

# review ratings
# NOTE: Stef or Timo, this doesnt take into account repeated values per team
review_mean <- msa_models %>% 
  summarise(mean_prating = mean(phon_rating, na.rm = TRUE),
            sd_prating = sd(phon_rating, na.rm = TRUE),
            mean_srating = mean(stat_rating, na.rm = TRUE),
            sd_srating = sd(stat_rating, na.rm = TRUE))
```

@@

`r team_desc_list$Initial$n_teams$val %>% xfun::numbers_to_words() %>% stringr::str_to_sentence()` teams initially signed up to participate in the study.
`r (team_desc_list$Initial$n_teams$val - team_desc_list$Final$n_teams$val) %>% xfun::numbers_to_words() %>% stringr::str_to_sentence()` of the signed-up teams dropped out during the analysis phase.
One Team Coordinator, Wim van Dommelen (who agreed to their feedback being shared), gave the following reason: "The reason is that [...] the project's approach is too much top-down instead of bottom-up. [...] I would rather prefer to explore the collected utterances and run different types of acoustic analyses, which would then be followed by statistical evaluation of the data."
We find this contribution to clearly speak to the main objective of the current study: investigate researcher's degrees of freedom in the speech sciences.
Based on our personal experience with research in the field, it is common practice to test as many different types of models, using as many different types of measurements, to answer one research question.
The remaining `r team_desc_list$Final$n_teams$val` teams submitted their analysis by the established deadline.
These are the teams that are included in this study.
Table \@ref(tab:team-descriptives-table) provides a summary of the contributing teams and their analyses.

@@

<!-- Begin Table --> 
\vspace{0.2in}
```{r}
#| label: team-descriptives-table
#| echo: false
#| results: 'asis'
#| eval: true

# Round and format numbers to exactly N digits
specify_decimal <- function(x, k) {
  out <- trimws(format(round(x, k), nsmall = k))
  return(out)
}

# Team descriptives
tbl_team_desc <- msa_models %>% 
  select(animal, years_from_phd, prior_belief, phon_rating, stat_rating, 
         all_rating) %>% 
  group_by(animal) %>% 
  summarize(across(years_from_phd:all_rating, mean)) %>% 
  pivot_longer(cols = -animal, names_to = "metric", values_to = "vals") %>% 
  na.omit() %>% 
  group_by(metric) %>% 
  summarize(min = min(vals), max = max(vals), avg = mean(vals, na.rm = T), 
            sd = sd(vals, na.rm = T)) %>% 
  mutate(across(-c("metric", "avg", "sd"), round)) %>% 
  mutate(across(-c("metric", "min", "max"), specify_decimal, k = 2)) %>% 
  mutate(desc = "team_desc", 
    metric = fct_relevel(metric, "years_from_phd"), 
   `Mean (SD)` = glue::glue("{avg} ({sd})"), 
    Range = glue::glue("{min}--{max}"), 
    Metric = case_when(
      metric == "years_from_phd" ~ "Years since PhD", 
      metric == "stat_rating" ~ "Self-report statistical ability", 
      metric == "prior_belief" ~ "Prior belief in $H_{1}$", 
      metric == "phon_rating" ~ "Acoustic analysis rating", 
      metric == "all_rating" ~ "Overall peer review rating"), 
    label = glue::glue("{desc}_{metric}")) %>% 
  select(label, desc, Metric, Min. = min, Max = max, `Mean (SD)`, avg, sd, Range) %>% 
  add_row(label = " ", desc = "team_desc", Metric = " ", 
          Min. = NA, Max = NA, `Mean (SD)` = NA, avg = NA, sd = NA, Range = NA) %>% 
  mutate(across(-c("label", "desc", "Metric"), ~replace(., is.na(.), " ")))

# Model descriptives 
tbl_model_desc <- msa_models %>% 
  select(animal, n_predictors:n_random, n_models) %>% 
  group_by(animal) %>% 
  summarize(across(everything(), mean)) %>% 
  pivot_longer(cols = -animal, names_to = "metric", values_to = "vals") %>% 
  na.omit() %>% 
  group_by(metric) %>% 
  summarize(min = min(vals), max = max(vals), avg = mean(vals, na.rm = T), 
            sd = sd(vals, na.rm = T)) %>% 
  mutate(across(-c("metric", "avg", "sd"), round)) %>% 
  mutate(across(-c("metric", "min", "max"), specify_decimal, k = 2)) %>% 
  mutate(desc = "model_desc", 
    metric = fct_relevel(metric, "n_models"), 
    `Mean (SD)` = glue::glue("{avg} ({sd})"), 
    Range = glue::glue("{min}--{max}"), 
    Metric = case_when(
      metric == "n_coefficients" ~ "n. coefficients", 
      metric == "n_models" ~ "n. models", 
      metric == "n_predictors" ~ "n. predictors", 
      metric == "n_random" ~ "n. random effects", 
      metric == "n_random_int" ~ "n. random intercepts", 
      metric == "n_random_slo" ~ "n. random slopes", 
      metric == "n_random_smooth" ~ "n. random smooths",
      metric == "n_terms" ~ "n. terms"), 
    label = glue::glue("{desc}_{metric}")) %>% 
  select(label, desc, Metric, Min. = min, Max = max, `Mean (SD)`, avg, sd, Range) %>% 
  mutate(across(-c("desc", "Metric"), as.character))

# Combine in single tibble (for printing in prose and table)
tbl_team_model_desc <- bind_rows(tbl_team_desc, tbl_model_desc) %>% 
  mutate(desc = case_when(
    Metric != "Overall peer review rating" & desc == "team_desc" ~ " ", 
    Metric != "n. coefficients" & desc == "model_desc" ~ " ", 
    TRUE ~ desc), 
    desc = case_when(
      desc == "team_desc" ~ "Teams", 
      desc == "model_desc" ~ "Models", 
      TRUE ~ " ")
    )

tbl_team_model_desc %>% 
  select(Descriptives = desc, Metric, Range, `Mean (SD)`) %>% 
  knitr::kable(format = "latex", booktabs = T, escape = F, 
    linesep = "", align = c("l", "l", "c", "r"), 
    caption = "Descriptive statistics for teams and models included in the 
    meta-analysis (n = 30).",
    label = "team-descriptives-table"
    ) %>% 
  kableExtra::kable_styling(., font_size = 11)

```
\vspace{0.2in}
<!-- END: Table -->

@@

<!-- Found effect 
JVC: this code is ready to go, but not all the rows in the google sheet are 
filled out (approx. 40 missing)
--> 
```{r}
#| label: found-effect
#| echo: false

found_effect <- msa_models %>% 
  select(animal, model_n, found_effect) %>% 
  na.omit() %>% 
  mutate(effect = if_else(found_effect == "yes", 1, 0))

atleast_1_effect <- found_effect %>% 
  group_by(animal) %>% 
  summarize(total = sum(effect)) %>% 
  filter(total != 0)

# Overall N of analyses claiming an effect
n_effects <- sum(found_effect$effect)

# Percent of effects
perc_effects <- (n_effects / nrow(msa_models) * 100) %>% round(., 2)

# N of teams claiming an effect
n_atleast_1 <- nrow(atleast_1_effect)

# Percent of teams claiming an effect
perc_atleast_1 <- (n_atleast_1 / length(unique(msa_models$animal))) * 100
```

Based on statistical framework, type of model, distribution family, fixed terms, and not including random effects, there were a total of `r nrow(unique_model_specs)` different model specifications.
When considering both acoustic and statistical analysis, we have found a grand total of `r nrow(unique_acou_stat)`, i.e. each submitted analysis was unique.
Of the submitted analyses, `r n_effects` claimed to have found a statistically reliable effect (`r perc_effects`%). 
`r n_atleast_1 %>% xfun::numbers_to_words() %>% str_to_sentence()` teams out of the `r length(unique(msa_models$animal)) %>% xfun::numbers_to_words()` (`r perc_atleast_1`%) submitting analyses reported at least one found effect.

### 2022-09-19 RR_manuscript/Draft_RR.tex
update table, added section at the end of statistical analyses re. number of teams claiming an effect

@@

\phantom{.}

\begin{tabular}{llrrr}
\toprule
Descriptives & Metric & Min. & Max & Mean (SD)\\
\midrule
Teams & all rating & 39 & 88.67 & 69.95 (14.34)\\
 & Phon rating & 41.25 & 88.33 & 71.51 (13.92)\\
 & Prior belief of $H_{1}$ & 48.5 & 92 & 69.19 (11.33)\\
 & Statistical ability & 33 & 93.33 & 69.46 (16.52)\\
 & Years since PhD & -3.75 & 12.38 & 4.77 (4.16)\\
 &  &  &  & \\
Models & n. coefficients & 1 & 11 & 3.5 (2.75)\\
 & n. models & 1 & 16 & 3.67 (3.44)\\
 & n. predictors & 1 & 4.2 & 1.59 (0.85)\\
 & n. random effects & 0 & 10 & 2.47 (2.05)\\
 & n. random intercepts & 0 & 10 & 1.91 (1.77)\\
 & n. random slopes & 0 & 4 & 0.63 (1.07)\\
 & n. random smooths & 0 & 0 & 0 (0)\\
 & n. terms & 1 & 4.27 & 1.78 (1.08)\\
\bottomrule
\end{tabular}

Eighty-four teams initially signed up to participate in the study.
Fifty-four of the signed-up teams dropped out during the analysis phase.
One Team Coordinator, Wim van Dommelen (who agreed to their feedback being shared), gave the following reason: ``The reason is that {[}\ldots{]} the project's approach is too much top-down instead of bottom-up. {[}\ldots{]} I would rather prefer to explore the collected utterances and run different types of acoustic analyses, which would then be followed by statistical evaluation of the data.''
We find this contribution to clearly speak to the main objective of the current study: investigate researcher's degrees of freedom in the speech sciences.
Based on our personal experience with research in the field, it is common practice to test as many different types of models, using as many different types of measurements, to answer one research question.
The remaining 30 teams submitted their analysis by the established deadline.
These are the teams that are included in this study.

@@

@@

Based on statistical framework, type of model, distribution family, fixed terms, and not including random effects, there were a total of 47 different model specifications.
When considering both acoustic and statistical analysis, we have found a grand total of 104, i.e.~each submitted analysis was unique.

>>>>

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

Eighty-four teams initially signed up to participate in the study.
Fifty-four of the signed-up teams dropped out during the analysis phase.
One Team Coordinator, Wim van Dommelen (who agreed to their feedback being shared), gave the following reason: ``The reason is that {[}\ldots{]} the project's approach is too much top-down instead of bottom-up. {[}\ldots{]} I would rather prefer to explore the collected utterances and run different types of acoustic analyses, which would then be followed by statistical evaluation of the data.''
We find this contribution to clearly speak to the main objective of the current study: investigate researcher's degrees of freedom in the speech sciences.
Based on our personal experience with research in the field, it is common practice to test as many different types of models, using as many different types of measurements, to answer one research question.
The remaining 30 teams submitted their analysis by the established deadline.
These are the teams that are included in this study.
Table \ref{tab:team-descriptives-table} provides a summary of the contributing teams and their analyses.

@@

\vspace{0.2in}
\begin{table}

\caption{\label{tab:team-descriptives-table}Descriptive statistics for teams and models included in the 
    meta-analysis (n = 30).}
\centering
\fontsize{11}{13}\selectfont
\begin{tabular}[t]{llcr}
\toprule
Descriptives & Metric & Range & Mean (SD)\\
\midrule
Teams & Overall peer review rating & 39--89 & 69.95 (14.34)\\
 & Acoustic analysis rating & 41--88 & 71.51 (13.92)\\
 & Prior belief in $H_{1}$ & 48--92 & 69.19 (11.33)\\
 & Self-report statistical ability & 33--93 & 69.46 (16.52)\\
 & Years since PhD & -4--12 & 4.77 (4.16)\\
 &  &  & \\
Models & n. coefficients & 1--11 & 3.50 (2.75)\\
 & n. models & 1--16 & 3.67 (3.44)\\
 & n. predictors & 1--4 & 1.59 (0.85)\\
 & n. random effects & 0--10 & 2.47 (2.05)\\
 & n. random intercepts & 0--10 & 1.91 (1.77)\\
 & n. random slopes & 0--4 & 0.63 (1.07)\\
 & n. random smooths & 0--0 & 0.00 (0.00)\\
 & n. terms & 1--4 & 1.78 (1.08)\\
\bottomrule
\end{tabular}
\end{table}
\vspace{0.2in}

@@

Based on statistical framework, type of model, distribution family, fixed terms, and not including random effects, there were a total of 47 different model specifications.
When considering both acoustic and statistical analysis, we have found a grand total of 104, i.e.~each submitted analysis was unique.
Of the submitted analyses, 37 claimed to have found a statistically reliable effect (21.76\%).
Nine teams out of the thirty (30\%) submitting analyses reported at least one found effect.
